---
title: "EDS 230 - Assignment 2"
author: "Naomi Moraes, Leilanie Rubinstein"
date: "2025-04-13"
output: html_document
editor_options: 
  chunk_output_type: console
format:
  html:
    embed-resources: true
toc: true
warning: false
message: false
---

# Introduction

In this document we will attempt to conduct sensitivity analysis on the almond profit function - created for this assignment. We will incorporate the almond yield anomaly function - created for the prior assignment - in the context of the study, "Impacts of future climate change on California perennial crop yields" Lobell et. al, 2006. 

Yield anomalies are specifically defined as the percentage changes from a reference period's average yields. To find the profit, we used the equation 

$$ Profit = Revenue - Costs $$
We will assume that the entire yield produced will be sold at market, and that the costs of production are by acreage (according to: [Sample Costs to Produce an Orchard](https://coststudyfiles.ucdavis.edu/uploads/cs_public/67/b7/67b72c81-5ce0-4462-a396-fbb62ce8564e/2019sacvalleyalmonds.pdf) ). Thus, our equation can be further broken down as such:

$$ Profit = (Quantity Produced * Market Price) - (ProductionCosts * Acreage)$$

## Set-Up

```{r}
# Load libraries
library(tidyverse)
```

```{r}
# Import climate data
clim <- read_delim(here::here("data", "clim.txt"))
```

```{r}
# Source the function
source(here::here("R","summary_yield.R"))
source(here::here("R","almond_profit.R"))
source(here::here("R","almond_yield.R"))
```

```{r}
# Call the function
profit <- almond_profit(clim)
```

# Sensitivity Analysis with Two Parameters

Let's perform a sensitivity analysis on two key parameters:
1. `acres` - Number of acres in production
2. `production_cost` - Cost of production per acre (in USD)

```{r sensitivity}
# Create samples with reasonable ranges
nsamples <- 100

acres_samples <- runif(n = nsamples, min = 50, max = 200)
costs_samples <- runif(n = nsamples, min = 3500, max = 4500)

# Combine parameters for sampling
params <- expand.grid(acres = acres_samples, production_cost = costs_samples)

params_subset <- params[sample(nrow(params), 100), ]

# Run the model for each parameter combination
results <- pmap(params_subset, function(acres, production_cost) {
  almond_profit(
    clim = clim,
    market_price = 5000,
    production_cost = production_cost,
    acres = acres
  )
})

# Extract year and yield from each result
yield_anomalies <- map_df(results, function(res) {
  res %>% 
    select(year, yield) %>%  # Changed from yield_anomaly to yield
    mutate(run_id = cur_group_id())
}, .id = "run_id")

# Add parameter values to results
yield_anomalies <- bind_cols(
  yield_anomalies,
  params_subset[as.numeric(yield_anomalies$run_id), ]
)

# Calculate summary statistics for each year
yield_summary <- yield_anomalies %>%
  group_by(year) %>%
  summarise(
    mean_yield = mean(yield),
    median_yield = median(yield),
    min_yield = min(yield),
    max_yield = max(yield),
    sd_yield = sd(yield)
  )
```
